### 2022-07-14 ReviewerResponse.md

Reviewer responses
------------------

Reviewer 1:
-----------
The reviewer makes the following conjecture:
"The fact that the COVID-19 mortality can be accurately reproduced by the Gompertz curve does not seems to be very surprising taking into account that this dynamical dependence was initially developed to describe mortality curve and confirmed in this application by a variety of examples during almost two centuries"

While Gompertzian growth has been shown to exist in natural growth phenomena like cancer growth, plant leaf growth, etc., it has not previously been shown in the communicable disease paradigm before covid, e.g. at large spatial distances (e.g. across countries). This should indeed be surprising for the following reason: In smaller spatial scales, e.g. bacterial growth as the reviewer alludes to in Lavrova et al., the exponentially limiting growth factor is quite plausible. In the larger national scales, this exponential limiting factor is not plausible nor possible under the SIR/SIRD model, as we show in the introduction. The SIR/SIRD models imply a logistic growth curve, which is indeed not what we see. This conclusion is only possible due to our daily cadence of mortality data, not available for previous epidemics. In a weekly cadence, for example, the separation between the Gompertz and the Logistic is more tenuous and less convincing.

As we discuss in the paper and as has been discussed by references like Petroni et al., Gompertzian growth implies coherent or collaborating growth. In a communicable disease paradigm, we show that this collaboration can only result from infinite interactions and communication speed as exemplified in our proposed network in section 3.1. This is due to the constraint that people must be in contact to propagate the disease. In the reviewer's example from Lavrova, this constraint is not present and is indeed modeled as a ubiquitous oxygen concentration. This is a crucial difference that seems to have been missed by the reviewer.

Further regarding the reviewers comment:
"It also should be noted that the statement "a discrepancy between reality and the current theory of communicable diseases" does not seems to be correct because the classic (either SIR or SEIR)  theory of communicable diseases does not distinguish between "Removed" and "Recovered" persons as the variable "R". [...] Thus, a more motivated statement should be about developing a subdivision of an epidemiological model aimed at distinguishing between two outcomes."

This is exactly why we include a model that does distinguish between "Removed" and "Recovered" in the SIRD model and show that it leads to the logistic curve. Further, the premise of the paper is not to see what additional parameters must be added to the SIRD-like models to obtain Gompertzian growth, but rather to show that with only 2 parameters, under the environmental stressor model, the observations are accurately predicted. We do in fact leave it as an open question when we address potential networks that can lead to Gompertz, but do provide one such network in section 3. This seems much more "motivated" due to the parsimonious nature of the environmental stressor model, as opposed to adding to a model that already has more parameters and yet cannot accurately predict observations.

Regarding comment:
"Although the authors introduced SIRD model further in the text, the clear description of interplay about assumptions of both (SIR and SIRD) approaches should be discussed with the premises of both."

As we already discuss the SIRD and the SIR model based on the same premises, this comment is not clear. We introduce the "Dead" or "Removed" group for the exact reasons outlined by the reviewer himself above. A further elaboration of what kind of "premises" should be discussed is unclear.

"In addition, it should be pointed out that Ref. [5] already references the work Anastassopoulou at al. (2020). Data-based analysis, modelling and forecasting of the COVID-19 outbreak. PloS One, 15, e0230405 as that where SIRD model was introduced." 
It is not clear how this is relevant to the manuscript. We chose to follow Carletti due to the clarity of their exposition, not because they were the first ones to use the SIRD model.

Our modeling framework of the Gompertz does of course have wide implications with regards to the mechanisms of growth: We postulate that growth is taking place with instant communication between units. In the case of plant photosynthesis this is not so far fetched considering that electrons in the electron transport chain are already shown to be non-local, i.e. wavelike rather than particle like.  See for example G. Panitchayangkoon et al. "Long-lived quantum coherence in photosynthetic complexes at physiological temperature" PNAS 2010. This of course breaks down the chemical paradigm where molecules are localized in regular biological processes.

Although it was not the wish of the authors to present too much observational data that does not fit with the current virus-theory of covid, but rather focus on a mathematical model that does indeed explain this simple set of daily mortality data with a simple 2-parameter model, it may be illuminating to the reviewers and editor to note that many animals have been observed to die at anomalous rates during the spring of 2020.

- Fischer L, Peters M, Merbach S, Eydner M, Kuczka A, Lambertz J, Kum- merfeld M, Kahnt K, Weiss A, Petersen H. Increased mortality in wild tits in north rhine-westphalia (germany) in 2020 with a special focus on suttonella ornithocola and other infectious pathogens. European Journal of Wildlife Research 2021;67(3). doi:10.1007/s10344-021-01500-7.
- Peris ́e-Barrios AJ, Tomeo-Mart ́ın BD, G ́omez-Ochoa P, Delgado-Bonet P, Plaza P, Palau-Concejo P, Gonz ́alez J, Ortiz-D ́ıez G, Mel ́endez- Lazo A, Gentil M, Garc ́ıa-Castro J, Barbero-Fern ́andez A. Humoral responses to SARS-CoV-2 by healthy and sick dogs during the COVID- 19 pandemic in spain. Veterinary Research 2021;52(1). doi:10.1186/ s13567-021-00897-y. 
- Duff P, Fenemore C, Holmes P, Hopkins B, Jones J, He M, Everest D, Rocchi M. Disease surveillance in england and wales, december 2019. Vet- erinary Record 2020;186(1):15–6. doi:10.1136/vr.m68
- HuB,WeiH,FanZ,SongY,ChenM,QiuR,ZhuW,XuW,XueJ,Wang F. Emergence of rabbit haemorrhagic disease virus 2 in china in 2020. Veterinary Medicine and Science 2020;7(1):236–9. doi:10.1002/vms3.332.
- Fukui H, Shimoda H, Kadekaru S, Henmi C, Une Y. Rabbit hemorrhagic disease virus type 2 epidemic in a rabbit colony in japan. Journal of Vet- erinary Medical Science 2021;83(5):841–5. doi:10.1292/jvms.21-0007.
- van Aarde RJ, Pimm SL, Guldemond R, Huang R, Mar ́e C. The 2020 elephant die-off in botswana. PeerJ 2021;9:e10686. doi:10.7717/peerj. 10686.
- Kambayashi Y, Bannai H, Tsujimura K, Hirama A, Ohta M, Nemoto M. Outbreak of equine coronavirus infection among riding horses in tokyo, japan. Comparative Immunology, Microbiology and Infectious Diseases 2021;77:101668. doi:10.1016/j.cimid.2021.101668.
- King S, Rajko-Nenow P, Ashby M, Frost L, Carpenter S, Batten C. Out- break of african horse sickness in thailand, 2020. Transboundary and Emerg- ing Diseases 2020;doi:10.1111/tbed.13701.
- Lu G, Pan J, Ou J, Shao R, Hu X, Wang C, Li S. African horse sickness: Its emergence in thailand and potential threat to other asian countries. Transboundary and Emerging Diseases 2020;doi:10.1111/tbed.13625.
- Castillo-Olivares J. African horse sickness in thailand: Challenges of controlling an outbreak by vaccination. Equine Veterinary Journal 2020;53(1):9–14. doi:10.1111/evj.13353.

Taken as a whole, this cannot be explained by a virus. Indeed it would require the "virus" to spread to birds across the world simultaneously which seems quite implausible. A more plausible mechanism is an environmental stressor caused by a planetary change in the environment.

Minor text issues:
 (page 8): the sentence "where the bra-ket notation" could be misleading because it induces associations with quantum mechanical terms, inner products, etc. Here angle brackets denote simply standard statistical averaging.
 Noted, this will be changed to a simple expectation symbol "E[.]"

(page 9): what does it mean "planetary perturbation"?
This is the postulated cause of the environmental stressor. We can elaborate on this in the manuscript.

Reviewer 2:
-----------
This review is difficult to take seriously as it seems the reviewer has not understood the concepts introduced in the paper, as exemplified by
1. A lack of a reasoned discussion of why the paper should not be published.
2. The use of words like "esoteric". This word implied a something "mysterious" that cannot be explained, even though thorough explanations and even examples in section 3 are given.
3. The use of the word "kind of" implies a lack of understanding.
4. The reference given as Rodrigues et al. does not address the reason why Gompertz is observed, but merely shows that the observations can be fit with such a model, as has been cited by multiple papers in the manuscript. This reference in particular is a far cry from explaining the mechanistic underpinnings of the Gompertz model including its unification of the communicable disease paradigm based on a network of infected and susceptible people, thus obviating the statement "The present manuscript does not have an additional value." which also seems to confuse the reviewers value system with that of the general scientific community.
5. No discussion of the technical aspects of the manuscript.

Taken together it is clear that reviewer 2 did not engage with the material at the sufficient level to understand it.

To the Editor
-------------
I thank you for considering this manuscript for publication, but I must say that the quality of the reviews was not what I expected. This paper has the potential to have large ramifications in the scientific community as it not only postulates an alternative interpretation of observations with regards to covid, but also offers an alternative theory for Gompertzian growth in general and will thus affect our understanding of cancer growth, plant growth, etc.

I ask if you could suggest this paper to a few other reviewers that will hopefully engage more authentically with the material. Especially reviewer 2 does not seem to have the level of competence for this kind of job. Reviewer 1 (which I suspect is Postnikov) has some reasoned points, but are based on a misunderstanding in my view, and thus render me without tangible corrections I can make to improve the delivery of the manuscript's message.